---
title: "STAT 1293 - Assignment 5"
author: "Gordon Lu"
date: "7/26/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: Gastric bypass surgery (15 points)
### 1a) Are the conditions for the use of the large sample procedures met? Explain. (2 points)

### Solution:
Yes. First observing that our data comes from a SRS, we can assume the outcomes are unbiased and random. Additionally, notice how each individual's response can be summarized with a binary outcome. That being, either "Receiving gastic bypass surgery and maintaining at least a 20% loss six years after surgery" or "Receiving gastic bypass surgery and not maintaining at least a 20% loss six years after surgery". Then, note how the response of one individual does not impact the response of another, thus individuals are independent. Lastly, note that `np` $\geq 10 = 418(0.76)\approx 318$ and $n(1-p) \geq 10 = 418(0.24) \approx 100$. Thus, we can conclude that the sample size if sufficiently large, and importantly that the sampling distribution of p is normally distributed.

### 1b) Give a 90% confidence interval for the proportion of those receiving gastric bypass surgery that maintained at least a 20% weight loss six years after surgery. (4 points)

### Solution:
```{r, eval = TRUE}
np <- 0.76*418
prop.test(np, 418, conf.level = 0.90, correct = F)$conf.int
```

### 1c) Does the study provide sufficient evidence to claim that more than 70% of those receiving gastric bypass surgery maintained at least a 20% weight loss six years after surgery? Use $\alpha = 0.05$ and choose correct=F. (9 points)

### Solution:
```{r, eval = TRUE}
prop.test(np, 418, 0.70, conf.level = 0.95, correct = F, alternative = "greater")
```

Conclusion: Yes, since p < 0.05, there is sufficient evidence to claim that more than 70% of those receiving gastic bypass surgery maintained at least a 20% weight loss six years after surgery.

## Problem 2: I refuse! (15 points)
### 2a) What are the sample proportions of offers being rejected for the two groups? Are they similar? (6 points)

### Solution:
```{r, eval = TRUE}
human_reject <- 18/38
computer_reject <- 6/38
human_reject; computer_reject
```

The sample propertions of offers being rejected for the human group is: $\frac{18}{38} \approx 0.47$ and for the computer group is: $\frac{6}{38} \approx 0.16$. So, it appears that the sample proportions are not similar.

### 2b) We suspect that emotion will lead to offers from another person being rejected more often than offers from an impersonal computer. Do a test to assess the evidence for this conjecture. Use \$alpha = 0.01$ and correct=T. (9 points)

### Solution:
```{r, eval = TRUE}
prop.test(c(18, 6), c(38,38), conf.level = 0.99, correct = T, alternative = "greater")
```

Decision: Since the p-value is 0.003319, we reject the null at the significance level `0.01`, and conclude that there is sufficient evidence to conclude that offers from another person is significantly larger than from an impersonal computer.

Conclusion: In other words, we have sufficient evidence to conclude that emotion will lead to offers from another person being rejected more than offers from an impersonal computer.

## Problem 3: Where do young people live? (30 points)
### 3a) Based on the data, is there evidence that the proportions of living at home among the four age groups are different? Conduct a chi-squared test at $\alpha = 0.05$. (12 points)

### Solution:
```{r, eval = TRUE}
matA <- matrix(c(324, 216, 378, 388, 337, 464, 318, 559), 2, 4)
rownames(matA) <- c("Parents' home", "Other places")
colnames(matA) <- c(19, 20, 21, 22)
qchisq(0.95, 3)
chisq.test(matA, correct = F)
```

Decision: Since the $\chi^2 > \chi^{2}_{3, 0.95} \rightarrow 84.354 > 7.815$ we have sufficient evidence to reject the null hypothesis.

Conclusion: Therefore, we have sufficient evidence to conclude that the proportions of living at home among the four given age groups are different.

### 3b) Conduct a trend test to see if there is a linear trend in the proportions of living at home. (10 points)

### Solution:
```{r, eval = TRUE}
live_total <- margin.table(matA, 2)
home <- matA["Parents' home", ]
prop.trend.test(home, live_total)
```

Decision: Since the p-value is very small, p-value < 2.2e-16, we have sufficient evidence to reject the null hypothesis.

Conclusion: Therefore, we have sufficient evidence to conclude that there is a linear trend in the proportions of living at home.

### 3c) Create a plot the show the relationship between proportion of livng at home and age. (8 points)

### Solution:
```{r, eval = TRUE}
home_p <- home/live_total
plot(colnames(matA), home_p, pch = 18, cex = 2, col = 4, xlab = "Age", ylab = "Proportion living at home")
lines(colnames(matA), home_p, lty = 2, col = 2)
```

It appears that the relationship between the proprotion of living at home and age is linear.
