---
title: "STAT 1293 - Quiz 3"
author: "Gordon Lu"
date: "7/17/2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: Men of few words? (15 points)
### 1a) Create a histogram, boxplot, and a Q-Q plot of the data. Put them together using `par()` function. (6 points)

### Solution:
```{r, eval = TRUE}
talking <- read.table("C:/Users/gordo/Desktop/talking.txt", header = TRUE) #read in talking.txt
par(mfrow = c(1,3), pty = "s")
hist(talking$Words)
boxplot(talking$Words, horizontal = TRUE)
qqnorm(talking$Words)
qqline(talking$Words, col = 2)
```

### 1b) Examine the data. Are there any outliers? Is the distribution symmetric or skewed? (4 points)

### Solution:
```{r}
summary(talking$Words)
```

By making use of the 5-number summary, and the generated plots above, it is apparent that there are 2 upper outliers from the box plot, and the Q-Q plot. By using the `1.5 x IQR` rule, any outliers above `13869 + (1.5 x (13869 - 7868))` are considered upper outliers. In particular, any value above `22870.5` are considered otuliers. Therefore, the two upper outliers are `28408` and `37786`. However, there do not seem to be any lower outliers. Additionally, by looking at the histogram, the boxplot and the Q-Q plot, it is apparent that the distribution is skewed. This can also be noticed since the `Mean > Median`, which typically indicates a positively/right-skewed distribution.

### 1c) Do the data give convincing evidence that the mean number of words per day of men at this university differs from 7,000? Conduct a one-sample t test. Show your R output and the 4 steps of hypothesis test. Use significance level $\alpha =0.01$. (5 points)

### Solution:
```{r, eval = TRUE}
t.test(talking$Words, mu = 7000, conf.level = 0.99)
```

4-Step H.T.

Hypothesis: $H_{0}: \mu = 7000$ vs. $H_{a}: \mu \neq 7000$ 

Test statistic: $t_{0} = \frac{\overline{x} - \mu_{0}}{s/\sqrt{n}}$ = `3.145`

P-value: $P(T < t_0) = P(T < t_0)$ = `0.005332`

Conclusion: Reject $H_{0}$, since p-value < 0.05.

## Problem 2: Comparing two drugs. (15 points)
### 2a) Create a variable, diff, which is the difference between the two drugs (Reference-Generic). Show the vector. (3 points)

### Solution:
```{r, eval = TRUE}
drugs <- read.table("C:/Users/gordo/Desktop/drugs.txt", header = TRUE) #read in drugs.txt
diff <- drugs$Ref - drugs$Generic
diff
```

### 2b) Create a histogram and a boxplot of diff. (5 points)

### Solution:
```{r, eval = TRUE}
hist(diff)
boxplot(diff, horizontal = TRUE)
```

Yes, the distribution is positively skewed, this is apparent through the shape of the histogram, and the upper outlier that is `2353`.

Yes, there is a single outlier, 2353, this is apparent by looking at the boxplot.

Yes, it seems reasonable to use the `t` procedure, since based on the histogram, the data looks fairly normal. Additionally, the boxplot implies the shape, disregarding the upper outlier, that the distribution is rather symmetric.

### 3c) Conduct a one-sample t test. Use significance level $\alpha = 0.05$. (7 points)

### Solution:
```{r, eval = TRUE}
t.test(diff, conf.level = 0.95)
```

4-Step H.T.

Let d = (Reference - Generic)

Hypothesis: $H_{0}: \mu_d = 0$ vs. $H_{a}: \mu_d \neq 0$ 

Test statistic: $t_{0} = \frac{\overline{x} - \mu_{0}}{s/\sqrt{n}}$ = `-0.15455`

P-value: $P(T < t_0) = P(T < t_0)$ = `0.8788`

Conclusion: Fail to reject $H_{0}$, since p-value > 0.05.

## Problem 3: Student drinking (20 points)
### 3a) Compare the summary statistics of the female sample and male sample (use `summary()`). Which group has a higher mean? (4 points)

### Solution:
```{r, eval =  TRUE}
drinks <- read.table("C:/Users/gordo/Desktop/drinks.txt", header = TRUE) #read in drinks.txt
drinks.male <- drinks$Drinks[drinks$Sex == "M"]
drinks.female <- drinks$Drinks[drinks$Sex == "F"]

summary(drinks.male)
summary(drinks.female)
```

It appears that the `male` group has a higher mean.

### 3b) Create histograms for female students and male students. Make sure that they are on the same scale (same `xlim`). (4 points)

### Solution:
```{r, eval = TRUE}
par(mfrow = c(1,2), pty = "s")
hist(drinks.male, xlim = c(0, 20), ylim = c(0, 30))
hist(drinks.female, xlim = c(0, 20), ylim = c(0, 30))
```

### 3c) Create a side-by-side boxplot for female and male students. (4 points)

### Solution:
```{r, eval = TRUE}
boxplot(drinks.female, drinks.male, horizontal = TRUE)
```

### 3d) Conduct a general two-sample t test (without equal-variance assumption). Do male students drink more than female students, on average? (8 points)

### Solution:
```{r, eval = TRUE}
t.test(drinks.male, drinks.female, alternative = "greater")
```

4-Step H.T.

Hypothesis: $H_{0}: (\mu_{Male} - \mu_{Female}) = 0$ vs. $H_{a}: \mu_{Male} > \mu_{Female}$ 

Test statistic: $t_{0} = \frac{\overline{x} - \mu_{0}}{s/\sqrt{n}}$ = `5.1934`

P-value: $P(T < t_0) = P(T < t_0)$ = `3.807e-07`

Conclusion: Reject $H_{0}$, since p-value < 0.05.